### 2018-03-19 LCTES/lctes18-reviews-16-rebuttal.txt
(>) lctes

We would like to thank the reviewers for the thoughtful comments, questions,
and criticism.
We will accommodate all clarifications and answers in a final version of the
paper.
Each comment is addressed individually as follows.

References:

`[1]`: *Francisco Sant'Anna and others. 2013. Safe System-level Concurrency on
        Resource-Constrained Nodes. In Proceedings of ACM SenSys'13.*

`[2]`: *Adriano Branco and others. 2015. Terra: Flexibility and Safety in
        Wireless Sensor Networks. In ACM Transactions on Sensor Networks.*

Review #16A
===========

> I would appreciate getting a size estimate of the user code, runtime, driver,
> power management module as the cortex-m0 is very small in ROM/RAM.

All code is compiled together into a single binary image, but the runtime is in
the order of a few kilobytes (<5Kb).
Previous work `[1]` shows an overhead of around 10% for ROM and 5% for RAM for
complete applications in comparison to equivalent implementations in C (network
protocols, radio driver, etc.).

> I would also appreciate knowing the amount of power saved with this approach
> on some real life application.
> If so, may we get a bulk estimate of the amount of power saved?

Even though the examples in the paper use standby effectively, they are too
simple to draw any conclusions about power savings in real-world applications.
That said, our approach is optimal in the sense that the language ensures idle
states and the runtime can always figure out the most efficient sleeping mode
possible.
Nonetheless, this is a work in progress and we acknowledge the importance of a
quantitative evaluation in the conclusion of the paper as future work.
The evaluation will have to be quite extensive to cover typical passive/active
patterns in IoT applications.

Review #16B
===========

> It is not clear to me how the timing requirement can be met under this
> framework.

Timing requirements can be addressed in two levels of guarantees.
For soft real-time requirements, synchronous reactions (i.e., code outside
ISRs) are safer and more straightforward to use because they support rich
control abstractions (e.g., `par/or`), prevent race conditions, and are
guaranteed to terminate in bounded time.
However, since reactions run atomically and cannot interrupt each other, a
non-negligible delay in starting times is possible.
For more strict requirements, ISRs (i.e., code inside `async/isr`) are
asynchronous and can perform sensitive operations much faster (such as filling
a serial buffer) while synchronous code is active.

> Also, a quantitative analysis on the performance and energy for real
> applications is needed to show this work’s value in practice.

A quantitative analysis for real applications will be done in future work.
Please, see last comment in `Review #16A`.

Review #16C
===========

>    - no empirical evaluation or comparison to existing work
>
>    - no discussion of related work

Empirical evaluation and comparison to existing work will be done in future
work.
Please, see last comment in `Review #16A`.

>    - no evaluation of the overhead imposed by the use of the language 
>
>   its unclear to me that synchronous semantics have a clear adoption
>   advantage given the cost/overhead of its implementation

The overhead is very little.
Please, see first comment in `Review #16A`.

> - doesn't this place an extra burden on programmers and a level of expertise
>   beyond that novices to control at the program level?
>
> - Can we not do something similar and simpler via libraries and wrappers
>   using existing low level languages?

The burden to deal with standby is mostly on the device drivers, which *"are
write-once code that is typically packaged and distributed in a software
development kit (SDK)."* `[Introduction]`
By transferring the work from the applications to the language level, the idea
is that novice programmers never have to deal with standby explicitly.
In general-purpose languages, by exposing libraries and wrappers, the
programmers would still have to call them explicitly, which is exactly what our
proposal intends to avoid.

Furthermore, we show in Figure 2 that, when introducing concurrency, the
structure of the program remains sequential and amenable to inference of the
appropriate sleep mode.
In comparison with Arduino, whose main goal is to lower the entry barrier for
embedded development, Céu also preserves the sequential structure for
concurrent applications.

> - the intro over does it a bit in my opinion with the emphasis on the G20
>   action plan.

OK. We will also discuss the motivations from an academic perspective.

Review #16D
===========

> While it is possible to integrate this approach with CEU, <...>, it is not
> clear whether the approach can be extended to regular programming languages.

Céu preserves a sequential programming structure for complex concurrent
applications.
Please, see third comment in `Review #16C`.

> The benefit of this approach in real systems has not been evaluated.

The evaluations of the benefits in real systems will be done in future work.
Please, see last comment in `Review #16A`.

Review #16E
===========

> Cons:
>  - Expressiveness of the DSL in representing real IoT app/driver semantics
>    not demonstrated
>
> A major hurdle of a practical DSL is its expressiveness, <...>.
> I <...> would strongly encourage the authors to look for some real IoT
> scenarios to validate this design.

Although we have not evaluated the energy efficiency aspects (see last comment
in `Review #16A`), previous work `[1,2]` evaluates the expressiveness of Céu in
the context of Wireless Sensor Networks and discusses the development of
drivers, network protocols, and full applications in Céu.