### 2018-04-11 LCTES/lctes.tex

\documentclass[sigplan]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage{xspace}

\usepackage{listings}
\lstdefinelanguage{ceu}{%
  language=C,
  morekeywords={%
    and, uint32_t, code, async, isr,
    await, bool, call, class, data, define, deterministic, do, each, else,
    emit, end, escape, event, every, finalize, hor, implementation, in,
    input, interface, loop, min, native, new, nohold, not, or, output, par,
    pool, pure, return, signal, spawn, tag, then, this, traverse, until,
    var, watching, when, with},
}
\lstset{
  basicstyle=\ttfamily,
  captionpos=b,
  columns=flexible,
  commentstyle=\rmfamily\itshape,
  escapeinside={||},
  escapeinside={|},
  frame=tb,
  keepspaces=true,
  keywordstyle=\bfseries,
  language=ceu,
  mathescape=true,
  numbersep=4pt,
  numberstyle=\scriptsize,
  upquote=true,
}

\newcommand{\CEU}{\textsc{C\'{e}u}\xspace}
\newcommand{\code}[1] {{\small{\texttt{#1}}}}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
%\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El Paso, Texas USA}
\acmConference[LCTES'18]{ACM SIGPLAN/SIGBED}{June 2018}{Philadelphia, USA}
%\acmYear{1997}
%\copyrightyear{2016}

%\acmPrice{15.00}

%\acmBadgeL[http://ctuning.org/ae/ppopp2016.html]{ae-logo}
%\acmBadgeR[http://ctuning.org/ae/ppopp2016.html]{ae-logo}

\begin{document}
%\title{A Language-Based Approach for Standby in Embedded Systems}
\title{WIP: Transparent Standby for Low-Power, Resource-Constrained Embedded Systems}
\subtitle{A Programming Language-Based Approach}

%\titlenote{Produces the permission block, and copyright information}
%\subtitlenote{The full version of the author's guide is available as \texttt{acmart.pdf} document}

\author[F.~Sant'Anna]{Francisco Sant'Anna}
\email{francisco@ime.uerj.br}
\affiliation{Rio de Janeiro State University, Brazil}
%
\author[A.~Sztajnberg]{Alexandre Sztajnberg}
\email{alexszt@ime.uerj.br}
\affiliation{Rio de Janeiro State University, Brazil}
%
\author[A.L..~de~Moura]{Ana L\'ucia de Moura}
\email{amoura@inf.puc-rio.br}
\affiliation{PUC-Rio, Brazil}
%
\author[N.~Rodrigues]{Noemi Rodrigues}
\email{noemi@inf.puc-rio.br}
\affiliation{PUC-Rio, Brazil}

% The default list of authors is too long for headers.
%\renewcommand{\shortauthors}{B. Trovato et al.}

\begin{abstract}
Standby efficiency for connected devices is one of the priorities of the
\emph{G20's Energy Efficiency Action Plan}.
%
We propose transparent programming language mechanisms to enforce that
applications remain in the deepest standby modes for the longest periods of
time.
%
We extend the programming language \CEU with support for interrupt service
routines and with a simple power management runtime.
%
Based on these primitives, we also provide device drivers that allow
applications to take advantage of standby automatically.
%
Our approach relies on the synchronous semantics of the language which
guarantees that reactions to the environment always reach an idle state
amenable to standby.
%
In addition, we show that programs in \CEU can keep a sequential syntactic
structure, even when applications require non-trivial concurrent behavior, to
lower the barrier of adoption.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
    <concept>
        <concept_id>10010520.10010553.10010562.10010564</concept_id>
        <concept_desc>Computer systems organization~Embedded software</concept_desc>
    <concept_significance>500</concept_significance>
    </concept>
    <concept>
        <concept_id>10011007.10011006.10011041.10011048</concept_id>
        <concept_desc>Software and its engineering~Runtime environments</concept_desc>
        <concept_significance>300</concept_significance>
    </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded software}
\ccsdesc[300]{Software and its engineering~Runtime environments}

\keywords{Arduino, Concurrency, Embedded Systems, Esterel, IoT, Standby}

%\begin{teaserfigure}
  %\includegraphics[width=\textwidth]{sampleteaser}
  %\caption{This is a teaser}
  %\label{fig:teaser}
%\end{teaserfigure}

\maketitle

\input{body}

\bibliographystyle{ACM-Reference-Format}
\bibliography{my,standby,other}

\end{document}