### 2018-05-01 LCTES/lctes.tex
(>) LCTES: versão final

\documentclass[sigplan]{acmart}

% TODO-FULL
%\linespread{0.96}
\setlength{\belowcaptionskip}{-10pt}

\usepackage{booktabs} % For formal tables
\usepackage{xspace}
\usepackage{balance}

\usepackage{listings}
\lstdefinelanguage{ceu}{%
  language=C,
  morekeywords={%
    and, uint32_t, code, async, isr,
    await, bool, call, class, data, define, deterministic, do, each, else,
    emit, end, escape, event, every, finalize, hor, implementation, in,
    input, interface, loop, min, native, new, nohold, not, or, output, par,
    pool, pure, return, signal, spawn, tag, then, this, traverse, until,
    var, watching, when, with},
}
\lstset{
  basicstyle=\small\ttfamily,
  captionpos=b,
  columns=flexible,
  commentstyle=\rmfamily\itshape,
  escapeinside={||},
  escapeinside={|},
  frame=tb,
  keepspaces=true,
  keywordstyle=\bfseries,
  language=ceu,
  mathescape=true,
  numbersep=4pt,
  numberstyle=\scriptsize,
  upquote=true,
}

\newcommand{\CEU}{\textsc{C\'{e}u}\xspace}
\newcommand{\code}[1] {{\small{\texttt{#1}}}}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

%\acmBadgeL[http://ctuning.org/ae/ppopp2016.html]{ae-logo}
%\acmBadgeR[http://ctuning.org/ae/ppopp2016.html]{ae-logo}

%%% If you see 'ACMUNKNOWN' in the 'setcopyright' statement below,
%%% please first submit your publishing-rights agreement with ACM (follow link on submission page).
%%% Then please update our instructions page and copy-and-paste the NEW commands into your article.
%%% Please contact us in case of questions; allow up to 10 min for the system to propagate the information.
%%%
%%% The following is specific to LCTES'18 and the paper
%%% 'Transparent Standby for Low-Power, Resource-Constrained Embedded Systems: A Programming Language-Based Approach (Short WIP Paper)'
%%% by Francisco Sant'Anna, Alexandre Sztajnberg, and Noemi Rodrigues.
%%%
%\setcopyright{ACMUNKNOWN}
\setcopyright{acmcopyright}
\acmPrice{15.00}
\acmDOI{10.1145/3211332.3211337}
\acmYear{2018}
\copyrightyear{2018}
\acmISBN{978-1-4503-5803-3/18/06}
\acmConference[LCTES'18]{19th ACM SIGPLAN/SIGBED Conference on Languages, Compilers, and Tools for Embedded Systems}{June 19--20, 2018}{Philadelphia, PA, USA}

\begin{document}
%\title{A Language-Based Approach for Standby in Embedded Systems}
\title{WIP: Transparent Standby for Low-Power, Resource-Constrained Embedded Systems}
\subtitle{A Programming Language-Based Approach}

%\titlenote{Produces the permission block, and copyright information}
%\subtitlenote{The full version of the author's guide is available as \texttt{acmart.pdf} document}

\author[F.~Sant'Anna]{Francisco Sant'Anna}
\email{francisco@ime.uerj.br}
\affiliation{Rio de Janeiro State University, Brazil}
%
\author[A.~Sztajnberg]{Alexandre Sztajnberg}
\email{alexszt@ime.uerj.br}
\affiliation{Rio de Janeiro State University, Brazil}
%
\author[A.L..~de~Moura]{Ana L\'ucia de Moura}
\email{amoura@inf.puc-rio.br}
\affiliation{PUC-Rio, Brazil}
%
\author[N.~Rodrigues]{Noemi Rodrigues}
\email{noemi@inf.puc-rio.br}
\affiliation{PUC-Rio, Brazil}

% The default list of authors is too long for headers.
%\renewcommand{\shortauthors}{B. Trovato et al.}

\begin{abstract}
Standby efficiency for connected devices is one of the priorities of the
\emph{G20's Energy Efficiency Action Plan}.
%
We propose transparent programming language mechanisms to enforce that
applications remain in the deepest standby modes for the longest periods of
time.
%
We extend the programming language \CEU with support for interrupt service
routines and with a simple power management runtime.
%
Based on these primitives, we also provide device drivers that allow
applications to take advantage of standby automatically.
%
Our approach relies on the synchronous semantics of the language which
guarantees that reactions to the environment always reach an idle state
amenable to standby.
%
In addition, in order to lower the programming barrier of adoption, we show
that programs in \CEU can keep a sequential syntactic structure, even when
applications require non-trivial concurrent behavior.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
<ccs2012>
    <concept>
        <concept_id>10010520.10010553.10010562.10010564</concept_id>
        <concept_desc>Computer systems organization~Embedded software</concept_desc>
    <concept_significance>500</concept_significance>
    </concept>
    <concept>
        <concept_id>10011007.10011006.10011041.10011048</concept_id>
        <concept_desc>Software and its engineering~Runtime environments</concept_desc>
        <concept_significance>300</concept_significance>
    </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded software}
\ccsdesc[300]{Software and its engineering~Runtime environments}

\keywords{Arduino, Concurrency, Embedded Systems, Esterel, IoT, Standby}

%\begin{teaserfigure}
  %\includegraphics[width=\textwidth]{sampleteaser}
  %\caption{This is a teaser}
  %\label{fig:teaser}
%\end{teaserfigure}

\maketitle

\input{body}

\bibliographystyle{ACM-Reference-Format}
\bibliography{my,standby,other}

\end{document}